\textbf{Suggestions for improving invited speaker diversity to reflect trainee diversity}

### 2020-02-07 submission/manuscript.tex
format for submission, fix typos
finish revisions and format resubmission

\documentclass[10pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Policy should change to improve invited speaker diversity and reflect trainee diversity},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{\textbf{Policy should change to improve invited speaker diversity and
reflect trainee diversity}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

@@

Ada K. Hagan, Ph.D.\({^1\dagger}\), Rebecca M. Pollet, Ph.D.\({^1}\),
and Josie Libertucci, Ph.D.\({^2}\)

@@

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:akhagan@umich.edu}{\nolinkurl{akhagan@umich.edu}}

1. Department of Microbiology \& Immunology, University of Michigan, Ann
Arbor, Michigan

2. Farncombe Family Digestive Health Research Institute, Department of
Medicine, McMaster University, Hamilton, Ontario, Canada

@@

\subsection{Conflicts of Interest}\label{conflicts-of-interest}

@@

The biomedical sciences have a problem retaining white women and
underrepresented minorities in academia. Despite increases in the
representation of these groups in faculty candidate pools, they are
still underrepresented at the faculty level, particularly at the Full
Professor level. The lack of diverse individuals at the Full Professor
level contributes to the attrition of women and under-represented
minorities, as it confirms unconscious biases. The presence of
unconscious biases contribute to feelings of not belonging by trainees
and are amplified by visual representation of who is presented as the
``top scientist in their field''. Top scientists are not only defined by
the attainment of Full Professorships, but also through invited seminar
series. Invitations for faculty to present their research at other
university departments is highly valued offer that provides an
opportunity for collaborations and networking. However, if invited
speakers do not represent the demographics of current trainees, these
visual representations of successful scientists may contribute to
decreased attitudes of self-identification as a scientist, ultimately
resulting in trainees leaving the field or the academy. In this study,
we compare invited-speaker demographics to the current trainee
demographics in one microbiology and immunology department and find that
trainees are not proportionally represented by speakers invited to the
department. Our investigation prompted changes in policy for how invited
speakers are selected in the future to invite a more diverse group of
scientists. To facilitate this process, we developed a set of tips and a
web-based resource that allows scientists, committees, and moderators to
identify members of under-served groups. These resources can be easily
adapted by other fields or sub-fields to promote inclusion and diversity
at seminar series', conferences, and colloquia.

@@

Long-standing systemic bias, sexism, and racism have contributed to the
under-representation of many racial and ethnic groups, as well as women,
in science, technology, engineering, and math (STEM) fields (1--4).
Specifically, within the field of biomedical research in the United
States, the proportion of underrepresented minorities at the full
professor level has remained consistently low at 4\% (survey data taken
from the NIH from 2001 to 2013), compared to the U.S. population, which
is 32.3\% (5, 6). Similar discrepancies exist for women in biomedical
sciences as full professorships are currently held mostly by men (7, 8).
As demographics of faculty within the biomedical sciences remains skewed
towards Caucasian men, the demographics of trainees (graduate students
and postdocs) are becoming more diverse (5).

Policy changes are needed to support inclusion of all individuals,
particularly in the biomedical sciences(9). To increase retention of
historically under-represented minorities (HURM), non-Caucasian/non-HURM
(NCNH) individuals, and white women in biomedical fields, it is
important for trainees to have visual representations of themselves as
scientists. The importance of representation in retaining a diverse
group of individuals in STEM fields is supported by social role theory
(10). Individuals make inferences about characteristics that are needed
to be successful in a given role by examining individuals that most
occupy that role (10, 11). However, there is a lack of diverse
scientific experts in academia so underrepresented minorities are not
seeing adequate visual representations of themselves at the faculty
level. Therefore, trainees who do not see representation of themselves
in senior faculty positions, may decide that they do not possess the
characteristics that are required to succeed.

Invited seminar series are common within biomedical departments across
the United States (12). Usually, seminar series' consist of faculty
members selecting a scientist from another institution to visit their
university and present their research, as well as meet with other
faculty members and trainees. Named lectureships follow the same format
but are decided by committee and are considered more prestigious because
they are named in honor of prominent local scientists. These seminar
series and lectureships provide an opportunity for trainees to be
exposed to research outside of their department. Additionally, being an
invited speaker provides the scientist with an opportunity to make
future collaborations and build their own \emph{curriculum vitae} (CV).
Scientists invited to give seminars are widely regarded as successful
and the top in their field. Thus, if trainees are constantly being
exposed to ``the top scientist in their field'', according to social
role theory, it signals who is successful in that field. While some have
examined this issue by studying and promoting the inclusion of more
women speakers at conferences, how department speaker series compare to
the trainee diversity of that department is unknown (13--15).

In this study, we examine and compare the proportion of HURM, NCNH, and
women invited speakers to white men in the Department of Microbiology
and Immunology at the University of Michigan. Additionally, we compare
invited-speaker demographics to the current trainee demographics as a
means to gauge if trainee demographics are being represented accordingly
throughout the seminar series. Following our investigation, we proposed
a policy change to the Department of Microbiology and Immunology in how
invited speakers are selected as a means to promote inclusion in our
department and reduce unconscious bias. In order to facilitate inviting
a more diverse group of scientists, we developed a set of resources that
allow scientists, within the fields of microbiology and immunology, to
self-identify as having an under-represented or under-served identity
including: HURM, non-Caucasian/non-HURM, or a white woman. These
resources will promote inclusion and diversity by providing greater
representation of all scientists and will provide hosts an opportunity
to invite a more diverse group of scientists.

@@

Each academic year, each faculty member in the Department of
Microbiology and Immunology at the University of Michigan has the
opportunity to invite one speaker per year for a weekly seminar series.
Some of these seminar slots are dedicated to named lectureships, which
are decided by committee, and three trainee-invited speakers. We
analyzed the demographics of invited speakers and faculty hosts for five
academic years (Fall 2014 - Spring 2019), and compared them to the
current trainees when the data were analyzed (Spring 2019). Each speaker
was only counted once and those listed as departmental faculty members
or as a ``host'' at any point could not also be considered ``invited
speakers''. The list of faculty hosts was used as a proxy for faculty
demographics since as hosts, these faculty members are visible
representatives of the department. The trainees were identified using
departmental email lists that included masters students, doctoral
students, and post-doctoral fellows.

This is a retrospective study, thus speakers were not asked for their
identities at the time of visit. Instead we hand-coded proxy
demographics using personal knowledge, photos, and CVs. The presenting
gender of each individual was assigned using a binary system
(man/woman). Due to the low number of individuals in the study,
race/ethnicity demographics were split in three groups: Caucasian,
Historically Under-represented Minority (HURM), and
Non-Caucasian/Non-HURM (NCNH), each with a binary (yes/no) possibility.
Caucasian was assigned using the current U.S. Census definition where
those of Middle Eastern, European, and Russian descent are included.
HURM individuals were restricted to those with African-American,
Indigenous, Alaskan/Hawaiian Native, Latinx and/or Hispanic heritage.
All others were placed into the NCNH group. We recognize that our proxy
demographics are a limitation of the analysis and want to acknowledge
that biological sex (male/female) is not always equivalent to the gender
that an individual presents as (man/woman), which is also distinct from
the gender(s) that an individual self-identifies as. We also want to
acknowledge that there are many other identities that are not captured
in this limited analysis.

Data were analyzed and figures generated in R Statistical Software,
using relevant packages (16--28).

@@

Our analysis identified an over-representation of Caucasian individuals
as hosting faculty and invited speakers (80\% each), relative to the
proportion of Caucasian trainees, which was 55\% (Fig. 1B). We also
observed declines in the representation of HURM and NCNH faculty and
speakers relative to the trainees (Fig 1B). HURM trainees made up 11\%
of the department, on track with the 11\% of U.S. microbiology and
immunology doctorates awarded in 2017 (29). However, only 8.5\% of
invited speakers, and none of the hosting faculty, were HURM scientists.
NCNH trainees were 34\% of department students and postdocs (versus 22\%
of U.S. microbiology and immunology doctorates in 2017), but only 19\%
of hosting faculty and 10.5\% of invited speakers (29).

The more prestigious invited speaker lectureships were also dominated by
Caucasian scientists, who comprised 81.48\% of those awarded (Fig. 1C).
HURM and NCNH scientists were awarded 3 and 2 lectureships,
respectively. Because the intersection of identities can compound biases
and outcomes, we further examined the lectureships by gender and
race/ethnicity status (30). Caucasian men and women accounted for
44.44\% and 37.04\% of the lectureships, respectively. Just 18.52\% of
lectureships were held by non-Caucasian men while none were held by
non-Caucasian women (Fig. 1D).

@@

This study found that the proportion of HURM and NCNH invited speakers
were under-representative of the trainee populations for each group.
Additionally, within the last 5 years, no HURM or NCNH woman was awarded
a lectureship. This means that the department is not providing
non-Caucasian trainees with adequate representation of successful
scientists. Taking this into context of social role theory, by not
adequately representing the diversity of all trainees, the department is
not supporting an inclusive environment in terms of visual faculty
representation. We also found that the proportion of women as faculty
hosts and speakers in our study population is equivalent to global
estimates that 40\% of microbiologists are women, though women only
represent about 30\% of academic biomedical faculty (7, 31). Women are
also over-represented as graduate students and postdoctoral fellows in
this department. Overall, Caucasian scientists are over-represented as
host faculty and invited speakers, compared to their presence as
trainees, particularly when lectureships were considered.

Several papers have investigated the representation of women at
scientific conferences, however, we have only identified one that
focused on invited speakers at universities (12). In their study,
Nittrouer et al, examined 3,652 talks at 50 U.S. institutions in 2013 -
2014 and found that women faculty are less likely to be invited
speakers, despite similar acceptance rates (12). We have not been able
to identify any publications examining scientific speaker diversity
beyond gender. This seems to be the first, which is concerning since
conclusions drawn from gender-based studies are often framed, and
considered, to be applicable to other marginalized groups (e.g., HURM).
This is a flawed assumption. While there is no doubt some overlap, each
group remains marginalized due to a unique complex set of factors that
cannot always be solved by gender-based solutions. U.S. institutions,
such as the University of Michigan have a particular responsibility to
the historically suppressed populations included in our definition of
HURMs. We therefore implore U.S. institutions to apply this framing to
their discussions and research.

Departments have different processes and criteria for selecting invited
speakers, but it is a matter of pride to bring the best scientists
possible. It may be that the definition of ``best'' poses a problem to
under-represented and under-served groups (e.g., white women, HURM, and
Asian) who are held to stricter competency standards and report having
to work harder than white men to be perceived as legitimate scholars
(32, 33). Some departments only invite tenured faculty, which severely
limits the number of potential speakers who are white women or
non-Caucasian. Yet, another scenario is that pre-tenure faculty members
invite prestigious, tenured faculty in their field to network and secure
letters for their own tenure package. The increased burden of white
women and non-Caucasian scientists to prove competency decreases their
likelihood to be considered for either tenure or as possible source of
tenure letters.

Each underrepresented group in our cohort faces a complex set of
barriers to achieving faculty status. For instance, the decision to
invite a woman may also be negatively impacted by assumptions about
competency and dedication. The dedication of women who have children to
their work is perceived to be less than that of their colleagues, i.e.,
men who also have children (34--36). The perceived prioritization and
commitments of women to family over work may cause faculty to doubt
their acceptance of a speaking invitation, despite the prestigious
nature of these invitations and evidence that men and women accept at
similar rates (12, 37). As a result, the faculty member may invite a
different colleague who they feel is more likely to agree (and is a
man). Another large portion of our sample were the NCNH cohort, who are
predominately Asian/Asian American individuals. Although Asian
scientists are well-represented in the US scientific workforce, they
face significant bias and barriers to inclusion in society and academia
(38, 39). For instance, despite the higher employment rate of Asian
scientists, they were not well-represented in the more prestigious
lectureships.

While HURM and NCNH share some experiences, differences including
varying rates of hiring and tenure promotion mean unique considerations
are important for inclusion of each group (3). For instance, a major
barrier to inclusion of HURM faculty at similar proportions to HURM
trainees is the low transition rate of scientists from HURM backgrounds
to faculty positions and the associated low proportion of HURM faculty
(40). The proportion of HURM faculty at the Assistant and Associate
Professor level is currently higher than at Full Professor so it will be
difficult to increase speaker diversity if early-career researchers are
not being considered (41). Increased performance expectations and
patterns of exclusions are consistent themes in studies characterizing
the HURM faculty experience (42, 43). Therefore, inclusion of HURM
faculty in seminar series is likely essential to increasing the number
of HURM Associate and Full Professors. Even when HURM speaker rates
match the proportion of HURM faculty employment, HURM trainees will be
represented at a significantly higher proportion. Inclusion of HURM
faculty in these seminar series is just one aspect of larger
institutional change that is needed (44).

@@

In an attempt to promote inclusion within the Department of Microbiology
and Immunology at the University of Michigan, these data were presented
to faculty members and the department chair (Dr.~Mobley). Since trainee
demographics were not represented by the seminar speaker demographics
over the past 5 years, we proposed a policy change as to how seminar
speakers were being invited. One suggestion was to switch from
faculty-invited to lab-invited speakers in an attempt to allow trainees
to choose a speaker that best represented themselves (Table 1).

The implicit biases that affect perceptions of marginalized groups are
an issue, but we must acknowledge that it is not always possible to
identify members of historically under-served communities. For instance,
after data analysis, we learned that at least one speaker in our data
set should have been categorized as a HURM instead of Caucasian, but it
wasn't readily apparent from their internet presence or CV. This
limitation makes two important points: that perceived identity often
plays a larger role than self-identification, and that we need better
tools to identify members of marginalized groups. Another policy
suggestion is for departments to invite their speakers to spend time
discussing their personal journeys through science, in addition to their
scientific stories (Table 1). This would enable those who wish, to
discuss how their identity(ies) interacted with their careers. In
addition to these suggestions for policy change, we have created
resources that allow scientists to self-identify as under-served groups
and thus provide host faculty with more diverse choices (Table 1).

>>>>

\documentclass[10pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Suggestions for improving invited speaker diversity to reflect trainee diversity},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{\textbf{Suggestions for improving invited speaker diversity to reflect
trainee diversity}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

@@

Ada K. Hagan, Ph.D.\(^{1,2\dagger}\), Rebecca M. Pollet, Ph.D.\({^2}\),
and Josie Libertucci, Ph.D.\({^3}\)

@@

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:akhagan@alliancescc.com}{\nolinkurl{akhagan@alliancescc.com}}

1. Alliance SciComm \& Consulting, Linden, Michigan

2. Department of Microbiology \& Immunology, University of Michigan, Ann
Arbor, Michigan

3. Farncombe Family Digestive Health Research Institute, Department of
Medicine, McMaster University, Hamilton, Ontario, Canada

@@

\subsection{Conflict of Interest Notification
Page}\label{conflict-of-interest-notification-page}

@@

Within the field of biomedical research in the United States, the
proportion of underrepresented minorities at the full professor level
has remained consistently low, even though trainee demographics are
becoming more diverse. Underrepresented groups face a complex set of
barriers to achieving faculty status including imposter syndrome,
increased performance expectations, and patterns of exclusion.
Institutionalized racism and sexism have contributed to these barriers
and perpetuated policy that excludes underrepresented minorities. These
barriers can contribute to decreased feelings of belonging, which may
result in decreased retention of underrepresented minorities. Though
some universities have altered their hiring practices to increase the
number of underrepresented minorities in the applicant pool, these
changes have not been sufficient. Here we argue that departmental
invited seminar series can be used to provide trainees with scientific
role-models and increase their sense of belonging, while institutions
work towards more inclusive policy. In this study, we investigated the
demographics (gender and race) of invited seminar speakers over 5 years
to the Department of Microbiology and Immunology at the University of
Michigan. We also investigated current trainee demographics and compared
them to invited speaker demographics to gauge if our trainees are being
provided with representation of themselves. We found that invited
speaker demographics were skewed towards Caucasian men, and our trainee
demographics were not being represented. From these findings, we
proposed policy change within the department to address how speakers are
being invited with the goal of increasing speaker diversity to better
reflect trainee diversity. To facilitate this process, we developed a
set of suggestions and a web-based resource that allows scientists,
committees, and moderators to identify members of under-served groups.
These resources can be easily adapted by other fields or sub-fields to
promote inclusion and diversity at seminar series', conferences, and
colloquia.

@@

Long-standing systemic bias, sexism, and racism have contributed to the
under-representation of many racial and ethnic groups, as well as women,
in science, technology, engineering, and math (STEM) fields (1--4).
Specifically, within the field of biomedical research in the United
States, the proportion of underrepresented minorities at the full
professor level has remained consistently low at 4\% (survey data taken
from the NIH from 2001 to 2013) (5, 6). Similar discrepancies exist for
women in biomedical sciences as full professorships are currently held
mostly by men (7, 8). As demographics of faculty within the biomedical
sciences remains skewed towards Caucasian men, the demographics of
trainees (graduate students and postdocs) are becoming more diverse (5).

Policy changes are needed to support inclusion of all individuals,
particularly in the biomedical sciences, since underrepresented groups
face a complex set of barriers to achieving faculty status (9). For
example, the dedication of women, who have children, to their work is
perceived to be less than that of their colleagues, including men who
also have children (10--12). Historically under-represented minorities
in the United States (HURM), Asian/Asian-Americans, and women are all
held to stricter competency standards and report having to work harder
than Caucasian men to be perceived as legitimate scholars (13, 14).
Asian/Asian-Americans suffer from imposter syndrome at greater rates
than other marginalized groups and Asian women report a lack of sponsors
(15, 16). Increased performance expectations and patterns of exclusions
are consistent themes in studies characterizing the HURM faculty
experience (17, 18). While HURM and other marginalized groups share some
experiences, differences including varying rates of hiring and tenure
promotion mean that unique considerations are important for inclusion of
each group (3).

Here, we argue that in order to increase the retention of faculty from
marginalized groups at the professor level, universities should provide
trainees with visual representation of themselves as successful
scientists. Recent studies show that women in STEM benefit from women
role models through improved belonging and self-efficacy (19, 20).
Predictably, a lack of active inclusion decreases self-efficacy,
eventually resulting in decreased feelings of belonging for URMs (21).
However, as the demographics of trainees have become more diverse,
students who are not Caucasian men, are lacking role models.
Institutionalized racism and sexism (22), defined as policies, societal
norms, and ideologies that reinforce inequities, have played a large
role in access to, inclusion in, and hiring policies at U.S.
universities (23, 24). Accordingly, faculty from marginalized groups are
eliminated from the applicant pool and subsequent hires, leaving
university policies and practices to be predominantly created by
Caucasian men. Thus, institutionalized racism and sexism are perpetuated
(25, 26). Universities have begun adjusting hiring practices and
creating initiatives to address inequitable hiring practices, but have
had limited results (27). Considering that most faculties within the
U.S. are still skewed towards Caucaisan men, invited seminar series are
a possible tool to provide marginalized trainees with representation of
themselves as successful scientists.

Invited seminar series are common within biomedical departments across
the United States (28). Usually, seminar series' consist of faculty
members selecting a scientist from another institution to visit their
university and present their research, as well as meet with other
faculty members and trainees. Named lectureships follow the same format
but are decided by committee and are considered more prestigious because
they are named in honor of prominent local scientists. These seminar
series and lectureships provide an opportunity for trainees to be
exposed to research outside of their department. Additionally, being an
invited speaker provides the scientist with an opportunity to make
future collaborations and build their own curriculum vitae (CV).
Scientists invited to give seminars are widely regarded as successful
and the top in their field, providing an opportunity for trainees to be
exposed to successful scientists in that field. Some studies have
examined the addition of more women speakers at conferences to promote
inclusion (29--31), however, we have only identified one other study
that has focused on invited speakers at universities (28). In their
study, Nittrouer et al, examined 3,652 talks at 50 U.S. institutions in
2013 - 2014 and found that women faculty are less likely to be invited
speakers (28). We have not been able to identify any publications
examining scientific speaker diversity beyond gender or how department
speaker series compares to trainee diversity in that department.

In this study, we examine and compare the demographics of invited
speakers to Caucasian men in the Department of Microbiology and
Immunology at the University of Michigan. Additionally, we compare
invited-speaker demographics to the current trainee demographics to
gauge if trainee demographics are being represented throughout the
seminar series. Following our investigation, we proposed a policy change
to the Department of Microbiology and Immunology in how invited speakers
are selected to promote inclusion in our department and reduce
unconscious bias. In order to facilitate inviting a more diverse group
of scientists, we developed a set of resources that allow scientists,
within the fields of microbiology and immunology, to self-identify as
having an under-represented or under-served identity including: HURM,
non-Caucasian/non-HURM (NCNH), or a Caucasian woman. These resources
will promote inclusion and diversity by providing greater representation
of all scientists and will provide hosts an opportunity to invite a more
diverse group of scientists.

@@

Each academic year, each faculty member in the Department of
Microbiology and Immunology at the University of Michigan has the
opportunity to invite one speaker per year for a weekly seminar series.
Some of these seminar slots are dedicated to named lectureships, which
are decided by committee, and three trainee-invited speakers. We
analyzed the demographics of invited speakers and faculty hosts for five
academic years (Fall 2014 - Spring 2019), and compared them to the
current trainees when the data were analyzed (Spring 2019). Each speaker
was only counted once and those listed as departmental faculty members
or as a ``host'' at any point could not also be considered ``invited
speakers''. The list of faculty hosts was used as a proxy for faculty
demographics since as hosts, these faculty members are visible
representatives of the department. There were a total of \{\texttt{r}\}
invited speakers and paired hosts. The trainees were identified using
departmental email lists that included masters students, doctoral
students, and post-doctoral fellows. There were \{\texttt{r}\} students
and \{\texttt{r}\} post-doctoral fellows.

This is a retrospective study, thus speakers were not asked for their
identities at the time of visit. Instead we hand-coded proxy
demographics of the speakers, faculty hosts and trainees using first
names, publicly available photos, and CVs (32--36). Information from
CVs, such as undergraduate institutions (e.g., Puerto Rico vs mainland
U.S.) and activity in HURM groups (e.g., SACNAS) helped inform our
demographics by indicating identities that might be held by that
individual. Because these data were collected from publicly available
sources, this study was not submitted to an IRB for consideration. The
presenting gender of each individual was inferred using a binary system
(man/woman). Due to the low number of individuals in the study,
race/ethnicity demographics were only split into three groups:
Caucasian, Historically Under-represented Minority (HURM), and
Non-Caucasian/Non-HURM (NCNH), each with a binary (yes/no) possibility.
Caucasian was inferred using the current U.S. Census definition where
those of Middle Eastern, European, and Russian descent are included. URM
individuals include those of African-American, Indigenous American,
Alaskan/Hawaiian Native, Latinx and/or Hispanic heritage (20 U.S. Code
§\,1067k), we use the HURM designation to recognize the history of
enslavement and active oppression in the U.S.. The NCNH group
predominantly included Asian/Asian-Americans, but also African
immigrants (37).

Data were compiled and figures generated in R Statistical Software,
using relevant packages (38--50).

@@

Our analysis identified an over-representation of Caucasian individuals
as hosting faculty and invited speakers (80\% each), relative to the
proportion of Caucasian trainees, which was 55\% (Fig. 1B). We also
observed declines in the representation of HURM and NCNH faculty and
speakers relative to the trainees (Fig 1B). HURM trainees made up 11\%
of the department, on track with the 11\% of U.S. microbiology and
immunology doctorates awarded in 2017 (51). However, only 8.5\% of
invited speakers, and none of the hosting faculty, were HURM scientists.
NCNH trainees were 34\% of department students and postdocs (versus 22\%
of U.S. microbiology and immunology doctorates in 2017), but only 19\%
of hosting faculty and 10.5\% of invited speakers (51).

The more prestigious invited speaker lectureships were also dominated by
Caucasian scientists, who comprised 81.48\% of those awarded (Fig. 1C).
HURM and NCNH scientists were awarded 3 and 2 lectureships,
respectively. Because the intersection of identities can compound biases
and outcomes, we further examined the lectureships by gender and
race/ethnicity status (52). Caucasian men and women accounted for
44.44\% and 37.04\% of the lectureships, respectively. Just 18.52\% of
lectureships were held by non-Caucasian men while none were held by
non-Caucasian women (Fig. 1D).

@@

This study found that the proportion of HURM and NCNH invited speakers
were under-representative of the trainee populations for each group.
Additionally, within the last 5 years, no HURM or NCNH woman was awarded
a lectureship, despite that in 2017 non-Caucasians were 30\% of the
professoriate (53). This means that the department is not providing
non-Caucasian trainees with adequate representation of successful
scientists and failing to support an inclusive environment in terms of
visual faculty representation. We also found that the proportion of
women as faculty hosts and speakers in our study population is
equivalent to global estimates that 40\% of microbiologists are women,
though women only represent about 30\% of academic biomedical faculty
(7, 54). Women are also over-represented as graduate students and
postdoctoral fellows in this department. Overall, Caucasian scientists
are over-represented as host faculty and invited speakers, compared to
their presence as trainees, particularly when lectureships were
considered.

We have not been able to identify any publications examining scientific
speaker diversity beyond gender (28). This seems to be the first, which
is concerning since conclusions drawn from gender-based studies are
often framed, and considered, to be applicable to other marginalized
groups (e.g., HURM). For instance, that African-American and Caucasian
women benefit equally from the same policy change. This is a flawed
assumption (55). While there is no doubt some overlap, each group
remains marginalized due to a unique complex set of factors that cannot
always be solved by gender-based solutions. The historical exclusion of
HURMs by U.S. institutions means that they have a particular
responsibility to improving the academic experience for these
populations (23). We therefore call on U.S. institutions to apply
intersectional framing to their discussions and research.

Departments have different processes and criteria for selecting invited
speakers, but it is a matter of pride to bring the best scientists
possible. The barriers to achieving faculty status for HURMs, Caucasian
women, and NCNH may also impact their speaking invitations. For
instance, the perceived prioritization and commitments of women to
family over work may cause faculty to doubt their acceptance of a
speaking invitation, despite the prestigious nature of these invitations
and evidence that men and women accept at similar rates (28, 56). As a
result, the faculty member may invite a different colleague who they
feel is more likely to agree (and is a man). It may also be that the
definition of ``best'' poses a problem to under-represented and
under-served groups (e.g., women, HURM, and Asian) who are held to
stricter competency standards (13, 14). Some departments may only invite
tenured faculty, which severely limits the number of potential speakers
who are Caucasian women or non-Caucasian. Yet, another scenario is that
pre-tenure faculty members invite prestigious, tenured faculty in their
field to network and secure letters for their own tenure package. The
increased burden of women and non-Caucasian scientists to prove
competency decreases their likelihood to be considered for either tenure
or as possible source of tenure letters. In particular, the proportion
of HURM faculty at the Assistant and Associate Professor level is
currently higher than at Full Professor so it will be difficult to
increase speaker diversity if early-career researchers are not being
considered (57). Even when HURM speaker rates match the proportion of
HURM faculty employment, HURM trainees will be represented at a
significantly higher proportion. We argue that inclusion of marginalized
faculty in seminar series is an important factor to increasing their
representation among Associate and Full Professors. It is just one
aspect of larger institutional change that is needed, but one that will
benefit trainee experiences and the CVs of marginalized faculty (58).

We recognize that our proxy demographics are a limitation of the
analysis and want to acknowledge that biological sex (male/female) is
not always equivalent to the gender that an individual presents as
(man/woman), which is also distinct from the gender(s) that an
individual self-identifies as. We also want to acknowledge that there
are many other identities that are not captured in this limited
analysis. While our pilot study combined 5 years worth of seminars, our
n is still quite low and we did not have other departments for
comparison. Consequently, our results can not be generalized to other
departments, fields, or universities. Another limitation to looking at a
single department is that trainees may interact with faculty from many
other departments, depending on their research and interests. Therefore,
the individual experience of representation would vary by trainee.
Future research needs to consider multiple departments, universities,
and/or fields to bolster generalizability. There is a paucity of
research on speaker identities other than their gender, so this also
needs to be addressed in future studies. However, we caution that
representation is a shallow metric on which a single-minded focus can
cause more harm than help (25). We recommend that future research also
survey trainee and speaker experiences and trainee participation in
seminar series to better understand the dynamics at play.

@@

In an attempt to promote inclusion within the Department of Microbiology
and Immunology at the University of Michigan, these data were presented
to faculty members and the department chair. Since trainee demographics
were not represented by the seminar speaker demographics over the past 5
years, we proposed a policy change as to how seminar speakers were being
invited (Table 1). One suggestion was to switch from faculty-invited to
lab-invited speakers to allow trainees to choose a speaker that best
represented themselves. This is easy to implement as it does not change
the overall structure of the department's seminar series; however, for
this same reason it may not have a significant impact on the diversity
of invited speakers. Trainees may be pressured to invite the top
individuals in their sub-field or may be influenced by the same
unconscious biases as faculty members. This idea can be expanded by
increasing the number of trainee-invited speakers and varying the
trainee group that extend invitations, for instance by training program,
career interest, and/or trainees and faculty in an identity-affinity
group. Invitations from trainees are often seen as an honor by potential
seminar speakers and nominating as a group may decrease the pressure to
invite particular sub-fields or ranks.

We also used this opportunity to begin a conversation about the purpose
of seminar speakers. Seminar speakers are sometimes invited to highlight
their latest high-impact paper or to share the arc of discoveries they
have made over several years of their career. While both are certainly
worthwhile, there are other benefits to be gained from interacting with
seminar speakers such as how to apply new techniques and how research is
framed outside research-focused universities (Table 1). Thinking more
broadly about what material is valuable during a seminar series may lead
to more speakers from under-represented and under-served backgrounds, as
well as more diversity of career paths. For some institutions, these
suggestions represent more of a structural change to the departmental
seminar series as speakers focusing on techniques or from
non-research-intensive universities are using invited as part of
professional development series. If these changes are to be implemented,
many members of the department must agree to the value of including
these seminars in the main departmental seminar series and these
expectations must be clearly communicated by the leadership (Table 1).
Departmental leadership can also ask individuals and groups to consider
their own unconscious biases that may be impacting their speaker nominee
list to compact some of the barriers to inviting diverse speakers.
Presented with these ideas, several members of the Department of
Microbiology and Immunology at the University of Michigan expressed
interest in specific resources they might use to identify individuals
from diverse identities, careers, and institutions. One suggestion is to
use resources that are being organized by professional societies such as
the American Society for Microbiology (ASM) and the Society for
Advancing Chicanos/Hispanics \& Native Americans in Science (SACNAS)
(Table 1). We also chose to develop ``Diversify'' resources for the
microbiology and immunology fields that provide a list of scientists
from under-represented and under-served groups, that is not associated
with a specific professional society. More information on the type of
resources and how to establish a Diversify list is below. Using list
resources like those available from professional societies as well as
our ``Diversify'' resources are particularly useful as social science
research has shown that the human brain is much better at recognizing
and using information (such as a strong scientific speaker) from a list
than it is at recalling the same information from memory (59, 60).
